---
title: "Piece_3"
author: "Benjamin Egan"
date: "2025-02-21"
output: html_document
---


## Monte Carlo Simulation - Brown M&Ms

As part of my responsibility as a TA for the PSY 312 course, I am responsible for the execution of lab exercises designed to aid in comprehension of topics learned during lecture. To help students understand probability, sampling, and the creation of a distribution of sample means, we did the "M&M Lab."

In the M&M lab, students discovered that Mars, Inc. does not distribute the colors evenly within a bag of plain M&Ms. This depends upon which of two packaging plants (Cleveland or New Jersey) made the product.

The back half of the lab has students run through a Monte Carlo Simulation to create a distribution of sample means using only brown M&Ms. Students are asked to create three samples with 1, 9, and 25 observations. We then plot them on a graph and see how the distribution looks.

### The problem

This normally works since there are 18 students, but Workday made it so my lab has only 9. To counteract this, I went in an manually simulated the other 9 students. Doing this was a pain, and I wanted to write a script that would do this for me.


## The Solution

```{r setup, message = FALSE, echo = FALSE}
library(haven)
library(tidyverse)
library(ggplot2)
```

```{r read the data}
brown <- read_spss("Brown_M&Ms_2025.sav")
tibble(brown)
```


In the data set, each value represents the number of brown M&Ms in a packet. There are 300 "packets" in the data set. I created a histogram below.


```{r initial visualtion, message = FALSE}

brown %>%
  ggplot(aes(
    x = brownmms
  ))+
  geom_histogram(color = "black", fill = "tan")+
  theme_bw()+
  labs(
    x = "Number of brown M&Ms in a pack",
    y = NULL,
    title = "Histogram of Brown M&M population"
  )

mean(brown$brownmms)
sd(brown$brownmms)
```

Here we can see that the "population" of brown M&Ms is fairly spread out(sd = 4.96), and centered around 11.5.

### Monte Carlo simulation

The first step is to get a sample of 1.

```{r monte sample size 1}
set.seed(1234)
sample(x = brown$brownmms, size = 1)

#use replicate() to get multiple of them

```

By using the sample() function, I am able to randomly sample 1 packet from the 300. For replication purposes, I set the seed so I know that i will always get the same numbers. This time I got 10, representing the average number of brown M&Ms in a packet with a sample size of 1.

I can also change the size of my sample.

```{r monte sample size 9 and 25 }
set.seed(1234)
sample_9_once <- sample(x = brown$brownmms, size = 9)
sample_25_once <- sample(x = brown$brownmms, size = 25)

mean(sample_9_once)
mean(sample_25_once)


#use replicate() to get multiple of them

```

I replicated the samples but increased my sample sizes. In the lab, we asked students to find the mean of each sample and record it.

### Replication

It would be pointless to copy and paste the lines of code 27 times to simulate 9 students obtaining three samples. So instead, I used repeat() to do this for me.

```{r replication basic}
set.seed(1234)
number <- 1

repeat { 
sample(x = brown$brownmms, size = 1) %>%
  print()
   number <- number + 1
   if(number >3) { 
      break
   }} 

```

Here I was able to create a sample with size 1, and repeat this function three times, simulating one person. Since I don't need to distinguish between participants, I can do it 27 times, creating a list.

```{r replication for size 1}
set.seed(1234)
number_size1 <- 1

repeat { 
sample(x = brown$brownmms, size = 1) %>%
  print()
   number_size1 <- number_size1 + 1
   if(number_size1 >27) { 
      break
   }} 

```

Now we need to replicate this with sample sizes of 9 and 25. Copy and pasting the code while changing the code provides us with a problem. It's printing out each number in the sample and not providing us with the mean of the sample. 


```{r replication for size 9 wrong}
set.seed(1234)
number_size9_wrong <- 1
repeat { 
sample(x = brown$brownmms, size = 9) %>%
  print()
   number_size9_wrong <- number_size9_wrong + 1
   if(number_size9_wrong >27) { 
      break
   }} 

```

To fix this, I piped the sample function inside of mean() before piping into print.

```{r replication for size 9 correct}
set.seed(1234)
size9_repeat <- 1
repeat { 
sample(x = brown$brownmms, size = 9) %>%
    mean() %>%
  print()
   size9_repeat <- size9_repeat + 1
   if(size9_repeat > 27) { 
      break
   }} 

```

Finally I can apply what I've learned to the sample size of 25.


```{r replication for size 25}
set.seed(1234)
size25_repeat <- 1
repeat { 
sample(x = brown$brownmms, size = 25) %>%
    mean() %>%
  print()
   size25_repeat <- size25_repeat + 1
   if(size25_repeat > 27) { 
      break
   }} 

```